---
title: "XXXXXX"
author: "XXXXXX"
date: "XXXXXX"
output: 
  html_document:
    number_sections: true
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 2
    theme: journal
    code_folding: show
---

#Preparation

```{r prep}
library(tidyverse)
```


#Part 1
##Load the data

```{r load_data}
data <- read.csv("titanic.csv")
```

##Delete columns

Get rid of the following columns: 'Name', 'Ticket', 'PassengerId', 'Cabin'
```{r del_col}
data <- data %>%
  select(-Name, -Ticket, -PassengerId, -Cabin)
```

##Replace

Replace the ‘Sex column with a binary column for whether the passenger is ‘Male’
```{r replace}
data <- data %>%
  mutate(Sex = ifelse(Sex == "male",1,0))
```

##Count NAs

Show how many NA’s are in each column
```{r count_NA}
for (i in 1:ncol(data))
  {n <- sum(is.na(data[i]))
  print(paste("Colums", colnames(data)[i], "has", n ," NAs."))}
```

##NA review

Which Column has the most NA’s? How many does it have?
```{r NA_review}
x <- c()
for (i in 1:ncol(data))
  {n <- sum(is.na(data[i]))
  x[i] <- n }
print(paste(colnames(data)[which.max(x)],"has the most NA. It has",x[which.max(x)],"NAs."))
```


##Bar chart

Create a bar chart to show NA counts by column
```{r bar_chart}
a <- sapply(data, function(x) sum(is.na(x)))
b <- colnames(data)
data1 <- data.frame(x = b, y = a)
ggplot(data1) +
  geom_bar(aes(x,y), stat = "identity")
```

##Plot

Show/plot the distribution of ages for each gender. Is the average (mean) age of Males higher? Is the median higher?
```{r distribution}
ggplot() +
  geom_boxplot(data = data,mapping = aes(as.factor(Sex), Age)) +
  geom_point(data = data %>% group_by(Sex) %>% summarise(mean = mean(Age, na.rm = TRUE)), mapping = aes(as.factor(Sex),mean), color = "red") +
  labs(x = "Sex")
```

Male has the higher mean and median of ages.


##Average fare
Which port of origin (embarked column) had the highest average fare? Show/plot your results.
```{r avg_fare}
data %>%
  filter(Embarked != "") %>%
  group_by(Embarked) %>%
  summarise(avg_fare = mean(Fare)) %>%
  ggplot(aes(Embarked,avg_fare)) +
  geom_bar(stat = "identity")
```

C port has the highest average fare.


##Statistically significant difference
Is there a statistically significant difference in average fares between the genders?
```{r stat}
t.test(data$Age[data$Sex == "1"],
       data$Age[data$Sex == "0"])
```

Because P value is less than 0.05, there is a statistically significant difference in average fares between the genders.


#Part 2
##Load data frame

```{r load}
library(jsonlite)
data2 <- readLines("bitly_usagov_example.txt")
data2 <- stream_in(textConnection(data2))
```

##Data review

How many records are there?
```{r records}
data2 %>%
  filter(!is.na(a)) %>%
  summarise(n = n())
```

##Count NAs
In the City column (‘cy’), how many are NA?
```{r num_NA}
sum(is.na(data2$cy))
```


##Count countries
How many countries are presented in this data? 
```{r num_c}
data2 %>%
  filter(!is.na(c)) %>%
  distinct(c) %>%
  summarise(n = n()) 
```
71 countries are presented in this data.


Which country (‘c’ column) has the most records? How many does it have?
```{r max_c}
data2 %>%
  filter(!is.na(c)) %>%
  group_by(c) %>%
  summarise(n = n()) %>%
  top_n(1)
```
US has the most records and it has 2305 records.


##From Russia
How many records are from Russia?
```{r from_R}
data2 %>%
  filter(c == "RU") %>%
  summarise(n = n()) 
```


##No USA
How many records do NOT list USA
```{r no_USA}
data2 %>%
  filter(!is.na(c),
         c != "US") %>%
  summarise(n = n())
```


##Recocrds search
how many records have 'America' in the timezone, but Country not in US?
```{r num_US}
data2 %>%
  filter(str_detect(tz, "^America"),
         c != "US") %>%
  summarise(n = n())
```

##Contains cia.gov
How many records point to cia.gov?
```{r cia}
data2 %>%
  filter(str_detect(u, "cia.gov")) %>%
  summarise(n = n())
```

15 records point to cia.gov.

##Fill records
Fill empty records with “UNKNOWN” and NA’s with “MISSING”. Create a new csv. (Submit this csv with your R Notebook.)
```{r fill}
data2[is.na(data2)] <- "MISSING"
data2[data2 == ""] <- "UNKNOWN"
data2$ll <- as.character(data2$ll)
write.table(data2,"data2.csv", sep = ",", row.names = FALSE)
```


##Plot
Show the top 10 timezones (‘tz’ column) in a bar chart, with a legend, and properly labeled x,y axes and title.
```{r tz_bar}
data2 %>%
  filter(tz != "MISSING" & tz != "UNKNOWN") %>%
  group_by(tz) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  top_n(10) %>%
  ggplot(aes(tz,n, color = tz)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x = "Timezone",
       y = "Count",
       title = "Top 10 timezones",
       color = "Timezone")
```































