---
title: "Week 4 data Lab"
author: "XXXXXX"
date: "XXXXXX"
output: 
  html_document:
    number_sections: true
    df_print: paged
    theme: journal
    code_folding: show
---

#Exercise 1:

Set working directory/ path to prepare your computer to load the csv
```{r exercise_1_1}
library(dplyr)
library(ggplot2)
library(tidyverse)
setwd("~/Documents/Santa Clara University/R Programing 2506/project/")
```

Load the csv file
```{r exercise_1_2}
data <- read.csv("data lab data.csv")
```

#Exercise 2:

Check if the csv already have preset column and column names
```{r exercise_2_1}
names(data)
```

Check and print the total numbers of NA in the table and Nas form each column

First, clean up data
```{r clean_up_data}
data <- data %>%
  filter(row_number() != 1L) %>%
  select(-(X)) %>%
  rename(all_ma_1980 = ALL.METROPOLITAN.AREAS,
         all_ma_1990 = X.1,
         all_ma_2000 = X.2,
         sel_ma_1980 = SELECTED.METROPOLITAN.AREAS,
         sel_ma_1990 = X.3,
         sel_ma_2000 = X.4)
```

Then, continue to process data
```{r exercise_2_2}
print(sum(is.na(data)))
for (i in 2:ncol(data))
  {n <- sum(is.na(data[i]))
    print(paste("Column", i, "has", n, "NAs"))}
```

#Exercise 3:

Make five separate sub-tables from the main table
(dissimilar index, isolation index, delta index, absolute centralizing index and spatial proximity index)

```{r exercise_3}
dissim_data <- data %>%
  filter(between(row_number(), which(grepl("Dissimilarity Index", data$Index.and.Race.Ethnicity))+1, which(grepl("Isolation Index", data$Index.and.Race.Ethnicity))-1))

iso_data <- data %>%
  filter(between(row_number(), which(grepl("Isolation Index", data$Index.and.Race.Ethnicity))+1, which(grepl("Delta Index", data$Index.and.Race.Ethnicity))-1))

del_data <- data %>%
  filter(between(row_number(), which(grepl("Delta Index", data$Index.and.Race.Ethnicity))+1, which(grepl("Absolute Centralization Index", data$Index.and.Race.Ethnicity))-1))

ab_cen_data <- data %>%
  filter(between(row_number(), which(grepl("Absolute Centralization Index", data$Index.and.Race.Ethnicity))+1, which(grepl("Spatial Proximity Index", data$Index.and.Race.Ethnicity))-1))

spa_data <- data %>%
  filter(between(row_number(), which(grepl("Spatial Proximity Index", data$Index.and.Race.Ethnicity))+1, n()))
```

#Exercise 4:

Show basic summaries of the main table and all five sub tables

You only need to do the means for the dissimilar index, and delta index in the exercise
```{r exercise_4_dissim_data}
dissim_data %>%
  gather("all_ma_1980","all_ma_1990","all_ma_2000", key = "metro_year", value = "data1") %>%
  gather("sel_ma_1980","sel_ma_1990","sel_ma_2000", key = "select_year", value = "data2") %>%
  group_by(Index.and.Race.Ethnicity) %>%
  summarise(mean_metro = mean(data1, na.rm = TRUE),
            mean_sel = mean(data2, na.rm = TRUE))
```


```{r exercise_4_del_data}
del_data %>%
  gather("all_ma_1980","all_ma_1990","all_ma_2000", key = "metro_year", value = "data1") %>%
  gather("sel_ma_1980","sel_ma_1990","sel_ma_2000", key = "select_year", value = "data2") %>%
  group_by(Index.and.Race.Ethnicity) %>%
  summarise(mean_metro = mean(data1, na.rm = TRUE),
            mean_sel = mean(data2, na.rm = TRUE))
```

#Exercise 5:

Make a graphic summary of the dissimilar index, do not include small multiples, just make simple plots, bar graph or visual representation of the dissimilar index
```{r exercise_5_metro}
dissim_data_new <- del_data %>%
  gather("all_ma_1980","all_ma_1990","all_ma_2000", key = "metro_year", value = "data1") %>%
  gather("sel_ma_1980","sel_ma_1990","sel_ma_2000", key = "select_year", value = "data2")

ggplot(dissim_data_new) +
  geom_line(aes(metro_year, data1, group = Index.and.Race.Ethnicity, color = Index.and.Race.Ethnicity)) +
  theme(legend.position="bottom")


```

```{r exercise_5_select}
ggplot(dissim_data_new) +
  geom_line(aes(select_year, data2, group = Index.and.Race.Ethnicity, color = Index.and.Race.Ethnicity)) +
  theme(legend.position="bottom")
```













